MATH 245 old and unused content

### 2024-02-16 PDEs/math245_hw2.Rmd
math hw prob 8

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-13"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

## Done, feel OK: 1, 2, 5, 7
## Done, very low confidence: 4, 6
## Not started: 3, 8

@@

which we know has the solution $v(x,y)=f(3x-y)$, so
$$u_x-u_y=f(3x-y)$$

On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
Set $\eta=x+y$ and $\xi=x$. 
Then by the chain rule,
$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.

So
$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$

$$u_{\xi}=f(4\xi-\eta)$$
Now integrate with respect to $\xi$:
$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
where $F$ is the antiderivative of $f$ with respect to $\xi$.

Now convert back to our original variables:
$$u(x,y)=F(3x-y)+G(x+y)$$

@@

>>>>

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-15"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

## Done, feel OK: 1, 2, 5, 7
## Done, very low confidence: 4
## Redo: 6
## Not started: 3, 8

@@

which we know has the solution $v(x,y)=f(3x-y)$, so
$$u_x-u_y=f(3x-y)$$
Now we can reorder our equation as $$\left(\partial_{x}-\partial_{y}\right)
\left( \partial_{x}+3\partial_{y}\right)u=0$$
and set $w=\left( \partial_{x}+3\partial_{y}\right)u$

Then $$w_x-w_y=0 $$

Which we know has the solution $w(x,y)=g(x+y)$. So $u_x+3u_y=g(x+y)$, which gives us a system of two equations:

\begin{equation}
\begin{cases}
u_x-u_y=f(3x-y)
u_x+3u_y=g(x+y)
\end{cases}
\end{equation}

Subtract the first equation from the second:
$$4u_y=-f(3x-y)+g(x+y) $$

Now we can integrate with respect to $y$ to find that:
$$u(x,y)=F(3x-y)+G(x+y)$$
where $F$ is the antiderivative of $-f$ with respect to $y$ and $G$ is the antiderivative of $g$ with respect to $y$.

@@

Set $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$. Then

$$\frac{\partial u}{\partial x} = \frac{\partial u}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial u}{\partial\eta}\frac{\partial\eta}{\partial x} = \frac{\partial u}{\partial \xi}+\frac{\partial u}{\partial\eta}$$

$$\frac{\partial^2 u}{\partial x^2} = \frac{\partial }{\partial x} \left(\frac{\partial u}{\partial \xi}\right) +\frac{\partial }{\partial x} \left(\frac{\partial u}{\partial \eta}\right)$$

$$=\frac{\partial^2 u}{\partial \xi^2} \frac{\partial \xi}{\partial x}+
\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial \xi}{\partial x} +
\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial\eta}{\partial x}+
\frac{\partial^2 u}{\partial \eta^2}\frac{\partial \eta}{\partial x}  $$

Which we can simplify to

$$u_{xx}= u_{\xi \xi}+2 u_{\xi \eta}+u_{\eta \eta}$$

Now we do the same for $y$:

$$\frac{\partial u}{\partial y} = \frac{\partial u}{\partial \xi} \frac{\partial \xi}{\partial y}+\frac{\partial u}{\partial\eta}\frac{\partial\eta}{\partial y} = y\frac{\partial u}{\partial \xi}-y\frac{\partial u}{\partial\eta}$$

But since $\eta=x-\frac{y^2}{2}$ and $\xi=x+\frac{y^2}{2}$, we can rewrite $y$ as $\sqrt{\xi-\eta}$.

$$\frac{\partial u}{\partial y} = \sqrt{\xi-\eta} \left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$

Now using the product rule,
$$\frac{\partial^2 u}{\partial y^2} = \sqrt{\xi-\eta}\; \frac{\partial}{\partial y}\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)+\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$

Where the second term has been simplified because $\frac{\partial}{\partial y}$ of $\sqrt{\xi-\eta}$ is simply $\frac{\partial}{\partial y}y=1$.

Now we use the chain rule again for the first term:

Now using the product rule,
$$\frac{\partial^2 u}{\partial y^2} = \sqrt{\xi-\eta}\left(\frac{\partial^2 u}{\partial \xi^2} \frac{\partial \xi}{\partial y}+
\frac{\partial^2 u}{\partial\eta \partial \xi}\frac{\partial\eta}{\partial y} -
\frac{\partial^2 u}{\partial \eta^2}\frac{\partial \eta}{\partial y}-
\frac{\partial^2 u}{\partial \eta \partial \xi}\frac{\partial \xi}{\partial y}\right)
+
\left(\frac{\partial u}{\partial \xi}-\frac{\partial u}{\partial\eta}\right)$$

Which simplifies to

$$u_{yy}= \left(\xi-\eta \right)\left(u_{\xi \xi}+u_{\eta \eta} -2 u_{\xi \eta}\right) +u_\xi-u_\eta$$

This means $y^3u_{xx}-yu_{yy}+u_y=0$ becomes

$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}+2 u_{\xi \eta}+u_{\eta \eta}\right) - \left(\xi-\eta \right)^{\frac{1}{2}} \left[\left(\xi-\eta \right)\left(u_{\xi \xi}-2 u_{\xi \eta}+u_{\eta \eta} \right) +u_\xi-u_\eta \right]+\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)=0  $$

$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}+2 u_{\xi \eta}+
u_{\eta \eta}\right) - \left(\xi-\eta \right)^{\frac{3}{2}} \left(u_{\xi \xi}-2 u_{\xi \eta}+
u_{\eta \eta}\right)-
\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)+
\left(\xi-\eta \right)^{\frac{1}{2}} \left(u_\xi-u_\eta\right)=0  $$

$$\left(\xi-\eta \right)^{\frac{3}{2}} \left(4\, u_{\xi \eta}\right)=0$$

But by integrating with respect to both variables, we recognize $u_{\xi \eta}=0$ as having the solution

$$u(\xi, \eta)=h_1(\xi)+h_2(\eta),$$

or in our original coordinates,

$$u(x,y)=h_1\left(x+\frac{y^2}{2}\right)+h_2\left(x-\frac{y^2}{2}\right) $$

When $x=2$,

$f(x)=h_1(4)+h_2(2x-4)$

$z=2x-4, \quad x=\frac{z}{2} + 2$

$h_2(z)=f\left(\frac{z}{2}+2\right)-h_1(4)$

and
$g(x)=h_1(2x)+h_2(0)$

$h_1(z)=g\left(\frac{z}{2}\right)-h_2(0)$

$h_1+h_2=g\left(\frac{z}{2}\right)-h_2(0)+f(\frac{z}{2} + 2)-h_1(4)$

But $h_2(0)+h_1(4)$ is just $g(x)$ when $x=2$ (or equivalently, $f(x)$ when $x=2$),

so after replacing $z$ by $x+\frac{y^2}{2}$ and $x-\frac{y^2}{2}$, we get

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+f(2)$$
\end{tcolorbox}

### 2024-02-16 PDEs/math245_old.Rmd
PDE tears
math hw prob 8

>>>>

Since $u=\frac{w}{r}$, this means we have
\begin{equation}u(t,r)=
    \begin{cases} 
      \frac{1}{2cr} \int_{r-ct}^{r+ct}{s\;ds} & r-ct \geq -1, r+ct \leq 1\\
      \\
      \frac{1}{2cr} \int_{-1}^{1}{s\;ds} & r-ct < -1, r+ct > 1\\
      \\
      \frac{1}{2cr} \int_{-1}^{r+ct}{s\;ds} & r-ct < -1, r+ct \leq 1\\
      \\
      \frac{1}{2cr} \int_{r-ct}^{1}{s\;ds} & r-ct \geq -1, r+ct > 1\\

    \end{cases}       
\end{equation}

On (characteristic) lines with the slope $y=-x+c$, or $y+x=\text{constant}$, we must have $u_x-u_y=f(3x-y)=0$.
Set $\eta=x+y$ and $\xi=x$. 
Then by the chain rule,
$$u_x= u_{\eta}+u_{\xi}, \; u_y=u_{\eta}$$
And let's rewrite $y$ as $y=\eta-x=\eta-\xi$.

So
$$u_x-u_y=f(3x-y) \quad\longrightarrow\quad u_{\xi}=f(3\xi-\eta+\xi)$$

$$u_{\xi}=f(4\xi-\eta)$$
Now integrate with respect to $\xi$:
$$u(\eta, \xi)=F(4\xi-\eta)+G(\eta)$$
where $F$ is the antiderivative of $f$ with respect to $\xi$.
